Water, High-Altitude Condensates, and Methane Depletion in the Atmosphere of the Warm Neptune WASP-107\lowercase{b}

### 2017-08-13 ms.tex
intro

\begin{abstract}
We report a measurement of the near-infrared transmission spectrum of the warm Neptune exoplanet WASP-107b measured with the Hubble Space Telescope Wide Field Camera 3. We detecta water absorption feature (FIXME confidence). The feature is small amplitude, clouds. We combine an interior structure model of the planet with an atmospheric retrieval . We find an upper limit on the atmospheric metallicity of FIXME solar, a cloud deck . 
\end{abstract}

@@

\section{Introduction} \label{sec:intro}

@@

\subsection{Atmospheric Metallicity from Interior Structure Modeling}
Given WASP-107b's unusually low density, we sought to estimate the range of atmospheric metallicities that are consistent with the observed mass and radius.  We estimated the bulk metallicity of WASP-107b using the structure evolution modeling of \cite{thorngren16}.  These models assume a thermally inert heavy-element core with a convective envelope of additively mixed H/He \citep{saumon95} and heavy-element impurities.  The heavy elements were a 50-50 rock-ice mix from ANEOS \citep{thompson90}.  We evolved the planets in time using the atmospheric models of \cite{fortney07}.  The results are sensitive to assumptions about the stellar age, which is uncertain \citep[either $0.6\pm0.2$ to $8.3\pm4.3$ Gyr depending on model assumptions;][]{mocnik17}. We therefore ran two models, with uniform age priors of either $0.2-1.0$ or $1.0-13.8$ Gyr.  We used the published mass and radius estimates \citep[$0.12\pm0.01\,M_\mathrm{J}$, $0.94\pm0.02$;][]{anderson17}.  A more precise mass estimate is in preparation (B. Benneke, priv.\,comm.); however, we find that our results are not significantly affected when we vary the assumed mass by $\pm3\,\sigma$ from the published value because the dominant uncertainties are the stellar age and core mass.

>>>>

\begin{abstract}
%We report a measurement of the near-infrared transmission spectrum of the warm Neptune exoplanet WASP-107b measured with the Hubble Space Telescope Wide Field Camera 3. We detecta water absorption feature (FIXME confidence). The feature is small amplitude, clouds. We combine an interior structure model of the planet with an atmospheric retrieval . We find an upper limit on the atmospheric metallicity of FIXME solar, a cloud deck . 
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
The composition of a planet's atmosphere depends on where and how the planet formed. By measuring atmospheric metallicity and molecular abundance ratios, we can shed light on important aspects of the formation process such as location within the disk and the relative accretion rates of gas versus solids \citep[][]{oberg11, fortney13, madhusudhan14, alidib16, mordasini16, espinoza17}.

The warm Neptune WASP-107b is an intriguing target for atmosphere characterization for several reasons.  It is an intermediate size between ice giants and gas giants, with a mass similar to Neptune's and a radius close to Jupiter's \citep[$0.12\,M_\mathrm{Jup}$, $0.94\,R_\mathrm{Jup}$;][]{anderson17}. Studying the atmospheres of planets in this transition region will provide additional clues in the much-debated mystery of what stunts the growth of Neptune-size planets \citep[e.g][]{pollack96, dawson16, frelikh17}.

WASP-107b also has a relatively low equilibrium temperature (780 K) compared to most other exoplanets that are amenable to atmosphere characterization.  This results in a distinct atmospheric chemistry from other well-studied systems: at low temperatures, the dominant molecular reservoir for carbon transitions from carbon monoxide to methane \citep{moses13}.  Spectral features from both water and methane are accessible with current observing facilities, raising the possibility of a spectroscopic estimate of the carbon-to-oxygen ratio (C/O). Previous measurements of C/O have been challenging because they rely on photometry that covers absorption features from multiple molecules \citep[e.g.][]{madhusudhan11, line14, benneke15, kreidberg15b}. FIXME: maybe cut this.

In addition, WASP-107b is one of the best targets discovered to date for atmosphere characterization. Thanks to its large atmospheric scale height and small, bright host star, the expected signal-to-noise for the transmission spectrum is comparable to the best studied benchmarks in the field (e.g. HD\,209458b, HD\,189733b). The planet is also a feasible target for thermal emission measurements, with an expected secondary eclipse depth of $\sim0.1$\% in \emph{Spitzer}'s $4.5\,\mu$m channel. In this paper we report the first atmosphere characterization of WASP-107b: a near-infrared transmission spectrum measured with the \emph{Hubble Space Telescope} (\emph{HST}).

@@

\subsection{Atmospheric Metallicity from Interior Structure Modeling}
Given WASP-107b's unusually low density, we quantiatively explored the range of atmospheric metallicities that are consistent with the observed mass and radius.  We estimated the bulk metallicity of WASP-107b using the structure evolution modeling of \cite{thorngren16}.  These models assume a thermally inert heavy-element core with a convective envelope of additively mixed H/He \citep{saumon95} and heavy-element impurities.  The heavy elements were a 50-50 rock-ice mix from ANEOS \citep{thompson90}.  We evolved the planets in time using the atmospheric models of \cite{fortney07}.  The results are sensitive to assumptions about the stellar age, which is uncertain \citep[either $0.6\pm0.2$ to $8.3\pm4.3$ Gyr depending on model assumptions;][]{mocnik17}. We therefore ran two models, with uniform age priors of either $0.2-1.0$ or $1.0-13.8$ Gyr.  We used the published mass and radius estimates \citep[$0.12\pm0.01\,M_\mathrm{J}$, $0.94\pm0.02$;][]{anderson17}.  A more precise mass estimate is in preparation (B. Benneke, priv.\,comm.); however, we find that our results are not significantly affected when we vary the assumed mass by $\pm3\,\sigma$ from the published value because the dominant uncertainties are the stellar age and core mass.
